An Empirical Analysis of the Role of Energy in Economic Growth

### 2013-03-14 EconGrowth-R-Functions.Rnw
Now creating CDe grid graphs in both paper and presentation. Need to fix size of presentation graphs and scaling of presentation graph text.

#############################
# This function creates a series of graphs for Cobb-Douglas with energy functions
# showing various parameters as a function of gamma.
##
cdeGridGraphs <- function(countryAbbrev, energyType){
  gammaGrid <- seq(0.0, 0.95, by=0.05)
  data <- cdeGridData(countryAbbrev, energyType, gammaGrid)
  with(data, plot(gamma, SSE))
  with(data, plot(gamma, lambda))
  with(data, plot(gamma, alpha))
  with(data, plot(gamma, beta))
  with(data, plot(beta, SSE))
}

>>>>

#############################
# This function creates a series of graphs for Cobb-Douglas with energy functions
# showing various parameters as a function of gamma.
# countryAbbrev tells which country you want to view
# energyType tells which type of energy you want
# textScaling tells how you want to scale the text on the axes and other places.
##
cdeGridGraphs <- function(countryAbbrev, energyType, textScaling=1.0){
  gammaGrid <- seq(0.0, 0.95, by=0.05)
  data <- cdeGridData(countryAbbrev, energyType, gammaGrid)
  # Now melt the data so that it is in the correct form.
  data_molten <- melt(data, id="gamma")
  alphaBetaGammaRange <- c(0.0, 1.0)
  alphaBetaGammaTics <- c(0.0, 0.2, 0.4, 0.6, 0.8, 1.0)
  lambdaRange <- c(0.0, 0.004)
  lambdaTics <- c(0.0, 0.001, 0.002, 0.003, 0.004)
  sseRange <- c(0.0, 0.08)
  sseTics <- c(0.0, 0.02, 0.04, 0.06, 0.08)
  # Create a lattice graph that shows the four variables (lambda, alpha, beta, and SSE) vs. gamma.
  graph <- xyplot(value ~ gamma | variable, data=data_molten,
                  # 1=lambda, 2=alpha, 3=SSE, 4=beta
                  index.cond=list(c(3,1,2,4)), #orders the panels as SSE, lambda, alpha, beta.
                  strip = strip.custom(factor.levels=c("$\\lambda$", "$\\alpha$", "$SSE$", "$\\beta$"), 
                                       bg="transparent", # Sets background transparent to match the graph itself.
                                       par.strip.text=list(cex=textScaling) # Scales text in the strip.
                                       ),
                  as.table=TRUE, #indexing of panels starts in upper left and goes across rows.
                  pch = c(1), col.symbol = c("black"), #Controls symbol parameters
                  scales=list(cex=scaleTextSize * textScaling, #controls text size on scales.
                              tck=scaleTickSize, #controls tick mark length. < 0 for inside the graph.
                              x=list(relation="free",
                                     at=alphaBetaGammaTics),
                              y=list(relation="free",  #allows each axis to be different
                                     #specifies location for tics
                                     at=list(lambdaTics, alphaBetaGammaTics, sseTics, alphaBetaGammaTics))
                              ), 
                  ylim=list(lambdaRange, alphaBetaGammaRange, sseRange, alphaBetaGammaRange), #y axis limits
                  #axis labels and scaling
                  xlab=list(label="$\\gamma$", cex=textScaling), 
                  ylab=list(label="")
                  )
  return(graph)
}

### 2013-03-14 EconGrowth.Rnw
Now using melt to reshape the data for graphs of lambda, alpha, beta, and SSE vs. gamma for CDe parameter estimation. Creating a lattice graph of the same variables.
Now creating CDe grid graphs in both paper and presentation. Need to fix size of presentation graphs and scaling of presentation graph text.

The Cobb-Douglas with energy parameter estimation process is problematic. The following graphs show that the minimum of the squared-errors curve occurs at the point where $\beta \rightarrow 0.0$.

<<CDeGammaGraphsZA, eval=TRUE>>=
  cdeGridGraphs("ZA", "Q")
@

>>>>

The Cobb-Douglas with energy parameter estimation process is challenging. To demonstrate the characteristics of the parameter estimation process, we performed a grid search over values of $\gamma$, wherein we re-estimated values of $\lambda$, $\alpha$, and $\beta$ at each value of $\gamma$. Figure \ref{fig:CDeGammaGraphsZA} shows the sum of squared errors ($SSE$), $\lambda$, $\alpha$, and $\beta$ as a function of the fixed value of $\gamma$ for South Africa with thermal energy ($q$). The minimum $SSE$ occurs at a discontinuity caused by the constraint $\beta \geq 0.0$. The parameter estimation algorithm has difficulty recognizing the values of $\gamma$, $\alpha$, $\beta$, and $\lambda$ that produce minimum $SSE$, because the constrained model behaves very differently on either side of the discontinuity.

<<CDeGammaGraphsZA, fig.cap="Cobb-Douglas with thermal energy ($q$) grid search for minimum $SSE$ for South Africa.", fig.pos="H", eval=TRUE>>=
cdeGridGraphs("ZA", "Q")
@
